---
title: "M7_Practice"
author: "Danni Martin"
date: "12/09/2024"
format: pdf
editor: visual
---

## Time Series IV - ARIMA Model

This module introduces the ARIMA model. ARIMA models are suitable for analyzing and forecasting time series data with constant mean and variance over time (stationary). By capturing the autocorrelation structure in the data, these models are particularly helpful in understanding the dynamics and relationships within a time series.

SES models based on trend and seasonality of data, while ARIMA describes autocorrelations in data.

Time series with trend or seasonality are NOT stationary.

Time series with cyclic behavior ARE stationary.

## Preliminaries

**Key Words/Concepts:**

-   White Noise: sequence of random data points with no correlation.

-   Stationary: no trend and homoskedastic

-   Differencing: Compute differences between consecutive observations. Make non-stationary series stationary.

-   Homoskedasticity: constant/uniform variance

-   Heteroskedasticity: increasing variance

-   Box-Cox Transformation: Used to normalize variance.

-   Unit Root Tests: tests to determine if series is stationary.

-   Autocorrelation: summarizes relationships between a series and its lags.

-   Partial Autocorrelation: Similar to ACF but intervening lags are removed.

**Models:**

-   AR: Autoregressive. Forecast using linear combination of past values of the variable.

-   MA: Moving Averages. Forecast using past errors.

-   ARIMA: Autoregressive Integrated Moving Averages. Combines all three by considering its own past, removing trends, and accounting for past errors.

### White Noise Example
White noise: sequence of random data points with no correlation. Benchmark for time series that exhibit patterns or trends.

```{r message=FALSE, warning=FALSE}
# Generate white noise process.
# Simulate data - normal distribution with a mean of zero and constant variance.
library(fpp3)
set.seed(10)
wn<-tsibble(x=rnorm(100),period=seq(1:100),index=period)

```

```{r}
# Plot.
# Series behaves erratically but fluctuates around mean of 0 with sd of 1.
wn %>% autoplot(x) + theme_classic() + 
  labs(title="White Noise Process",
       subtitle="Mean=0 and Standard Deviation=1",
x="",y="") + 
  geom_hline(yintercept = 0, col="blue", lwd=1, linetype="dashed",
             alpha=0.4)
```

### Stationary & Differencing

Time series are stationary when statistical properties do not change over time, ie constant mean, variance, and auto-covariance. White noise process is an example of a stationary process.

Stationary: No trend and is homoskedastic.

Many time models, including ARIMA, require a stationary time series. ARIMA models do no work well if the statistical properties change over time.

```{r}
# Load dataset
tesla<-tsibble(
  period=yearquarter(c("2016:Q1","2016:Q2","2016:Q3","2016:Q4",
                       "2017:Q1","2017:Q2","2017:Q3","2017:Q4",
                       "2018:Q1","2018:Q2","2018:Q3","2018:Q4",
                       "2019:Q1","2019:Q2","2019:Q3","2019:Q4",
                       "2020:Q1","2020:Q2","2020:Q3","2020:Q4",
                       "2021:Q1","2021:Q2","2021:Q3","2021:Q4",
                       "2022:Q1","2022:Q2","2022:Q3","2022:Q4",
                       "2023:Q1","2023:Q2")),
  deliveries=c(14.8,14.4,24.5,22.2,
               25,22,26.2,29.9,
               30,40.7,83.5,90.7,
               63,95.2,97,112,
               88.4,90.7,139.3,180.6,
               184.82,201.25,241.3,308.6,
               310.5,254.7,343.8,405.3,
               422.9,466.1),
  index=period     # This is the time variable
)
```

```{r}
# Plot tesla deliveries.
# Upward trend; therefore, series is NOT stationary.
tesla %>% autoplot(.vars=deliveries) + theme_classic() +
  labs(title= "Tesla Car Deliveries", 
       subtitle = "Q1 2017 to Q2 2023") +
  xlab("Quarter") + ylab(" ") +
  geom_hline(yintercept = mean(tesla$deliveries), col="blue", linetype="dashed", lwd=1, alpha=0.4)
```

Differencing: compute differences in between consecutive observations. Used to make non-stationary series stationary.

```{r}
# First difference.
# Eliminate trend using the difference() function. See plot.
# Seems to be some correlation between quarters.
tesla %>%
  as_tsibble(index=period, regular=T) %>% autoplot(difference(deliveries)) + theme_classic() +
  labs(title="Change In Tesla's Vehicle Deliveries",
       subtitle = "Q1 2016 - Q4 2022",x="",y="") + 
  geom_hline(yintercept = mean(difference(tesla$deliveries), na.rm = TRUE), col="blue", linetype="dashed", lwd=1, alpha=0.4)
```

```{r}
# Normalize variance in the series using a Box-Cox transformation.
lambda <- tesla %>% as_tsibble(index=period, regular=T) %>%
  features(deliveries, features = guerrero) %>% pull(lambda_guerrero) 

tesla %>% as_tsibble(index=period, regular=T) %>%
  autoplot(box_cox(difference(deliveries), lambda)) +
  labs(y = "")+ theme_classic() +
  labs(title="Box-Cox Transformation of Tesla's Vehicle Deliveries",
       subtitle = "Q1 2016 - Q4 2022",x="",y="")
```

### Unit Root Tests

Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test In this test, null hypothesis is that data are stationary, looking for evidence that the null hypothesis is false. Small p values suggest null hypothesis is false; therefore, differencing is required.

ndiffs determines number of differences required to make data stationary.

```{r}
tesla %>% as_tsibble(index=period, regular=T) %>%
  features(deliveries, features = c(unitroot_kpss, unitroot_ndiffs)) 
```

### Autocorrelation

```{r}
# Income growth in California.
# Load data and create test/train sets.
library(fpp3)
library(tidyverse)
PI<-read_csv("https://jagelves.github.io/Data/PersonalIncome.csv")
PI %>% as_tsibble(index=Date) %>% 
  filter_index(1970~2005) -> PI_train
PI %>% as_tsibble(index=Date) %>% 
  filter_index(2006~2021) -> PI_test
```

```{r}
# Apply and plot ACF function. Blue line indicates which ACs are statistically significant.
# Correlation decreases as lag increases.
PI_train %>%
  ACF(lag_max = 12,PI_Growth) %>% 
  autoplot() + theme_bw() + labs(x="", y="",
                                 title="ACF Personal Income Growth in California")
```

```{r}
# Apply ACF to white noise data.
# No correlation.
wn %>% ACF(x) %>% autoplot() + theme_bw() + labs(x="", y="ACF") +
labs(x="", y="", title="ACF White Noise Process")
```

### Partial Autocorrelation Function

```{r}
# Apply and plot PACF.
# Correlation only with first lag. Latter lags due to first lag influence.
PI_train %>%
  PACF(lag_max = 12,PI_Growth) %>% 
  autoplot() + theme_bw() + labs(x="", y="PACF") +
labs(x="", y="", title="PACF Personal Income Growth In California")
```

```{r}
# Apply PACF to white noise data. Confirm no correlation.
wn %>% PACF(x) %>% autoplot() + theme_bw() + labs(x="", y="PACF") +
labs(x="", y="", title="PACF White Noise Process")
```

## AR(1) Model

Uses only the first lag of the series as independent variable.

AR(p) uses largest lag.

```{r}
# Simulate data using AR(1) model.
y<-c(0)
phi<-0.7
const<-1
nrep<-100

for (i in 2:nrep){
y[i]=const+phi*y[i-1]+rnorm(1,0,0.5)
}
```

```{r}
# Apply ACF.
tsibble(y=y,period=seq(1,length(y)),index=period) %>% ACF(lag_max = 12, y) %>% autoplot()+theme_bw()+
labs(x="", y="", title="ACF For Simulated AR(1) phi=0.7, c=1")
```

```{r}
# Apply PACF.
tsibble(y=y,period=seq(1,length(y)),index=period) %>% PACF(lag_max = 12, y) %>% autoplot()+theme_bw()+labs(x="",y="PACF",                         title="PACF For Simulated AR(1) phi=0.7, c=1")
```

### Modeling and Residuals

Model PI growth using AR(1) model. Include LS model to compare.

```{r}
PI_fit<-PI_train %>% 
  model(AR1 = AR(PI_Growth ~ order(1)),
        LS = TSLM(PI_Growth ~ trend()))
coef(PI_fit)
```

```{r}
# Inspect errors using augment() function.
errors_PI<-augment(PI_fit)

# Plot ACF error.
# ACF of errors for AR(1) model resemble white noise.
errors_PI %>% select(.resid) %>% ACF(.resid) %>% 
  autoplot() + theme_bw() +
  labs(title="ACF for Model Errors of AR(1) and LS",
       x="")
```

```{r}
# Plot PACF error to confirm.
errors_PI %>% select(.resid) %>% PACF(.resid) %>% 
  autoplot() + theme_bw()
```

### Model Selection

```{r}
# Accuracy measures for AR(1) and LS models.
# AR performs better on all metrics.
glance(PI_fit) %>% arrange(AICc) %>% select(.model:BIC)
```

```{r}
# Confirm with test set.
PI_fc<-PI_fit %>% forecast(new_data = PI_test)
PI_fc  %>% accuracy(PI_test) 
```

```{r}
# Plot test set along with AR(1) forecast.
# Blue line indicates mean of predictions.
PI_fc %>% filter(.model=="AR1") %>% autoplot(level=95) + theme_classic() +
  autolayer(PI_train, PI_Growth) +
  autolayer(PI_test, PI_Growth) + 
  labs(title="Personal Income Growth AR(1) Forecast Accuracy",
subtitle="1970-2021", y="",x="")
```

## MA(1) Model

Moving average (MA) process models time series by using past errors (random shocks).

As with the AR model, further error term lags could be included in the MA. An MA process with highest lag of is called an MA(q) process. The MA process assumes that the current observation of a time series is a linear combination of the weighted sum of past error terms. The error terms capture the unpredictable and random fluctuations in the time series that are not accounted for by the AR model.

The pattern shown is one significant spike in the first lag of the ACF, and a decaying PACF as lags get larger. This is the opposite of the AR(1) process where we had a decaying ACF and one spike in the PACF.

```{r}
# Simulate MA(1) process and generate ACF/PACF plots.
set.seed(13)
e_t<-rnorm(300,0,0.5)
y_t=2+0.95*dplyr::lag(e_t)+e_t

tsibble(y=y_t,period=seq(1,length(y_t)),index=period) %>% ACF(lag_max = 12, y) %>% autoplot()+theme_bw()+
labs(x="", y="", title="ACF For Simulated MA(1)")

tsibble(y=y_t,period=seq(1,length(y_t)),index=period) %>% PACF(lag_max = 12, y) %>% autoplot()+theme_bw()+
labs(x="", y="", title="PACF For Simulated MA(1)")
```

### US Treasury Example

```{r}
# Load data and plot series.
treasury<-read_csv("https://jagelves.github.io/Data/treasury.csv")

treasury %>% mutate(period=1:length(OBS)) -> treasury
as_tsibble(treasury, index = period) %>%
  autoplot() + theme_classic() + 
  labs(title="Percent Monthly Change the Yield of U.S. Treasury Securities",
       subtitle="1953-2008",
       x="", y="")
```

```{r}
#ACF and PACF
as_tsibble(treasury, index = period) %>% 
  ACF(lag_max = 12, changetreasury) %>% 
  autoplot()+ theme_bw() +
labs(x="", y="", title="ACF for TCM5Y")

as_tsibble(treasury, index = period) %>% 
  PACF(lag_max = 12, changetreasury) %>% 
  autoplot()+ theme_bw() +
labs(x="", y="", title="PACF for TCM5Y")
```

## ARIMA Model

The ARIMA model is based on stationary, AR, and MA processes. ARIMA stands for Autoregressive Integrated Moving Average.

### Chilango Restaurant Example

```{r}
# Load data and create test/train sets.
cali<-read_csv("https://jagelves.github.io/Data/CaliforniaAvocado.csv")

cali %>%
  as_tsibble(key=geography,index=date,regular=T) %>%
  filter_index("2015-01-04"~"2018-12-02") -> cali

cali %>%
  as_tsibble(key=geography,index=date,regular=T) %>%
  filter_index("2015-01-04"~"2018-06-02") -> calits_train

cali %>%
  as_tsibble(key=geography,
             index=date, regular=T) %>%
  filter_index("2018-06-02"~"2018-12-02") -> calits_test
```

```{r}
# Estimate models.
fit <- model(calits_train,ETS=ETS(average_price),
              ARIMA=ARIMA(average_price,approximation = F),
              AR=AR(average_price~order(1)),
              LS=TSLM(average_price~trend()))
```

```{r}
# Check coefficients.
fit %>% coef()
```

```{r}
calits_train %>% ACF(lag_max=12, average_price) %>%
autoplot()+theme_bw()+
labs(x="", y="", title="ACF For Avocado Average Price")

calits_train %>% PACF(lag_max=12, average_price) %>%
autoplot()+theme_bw()+
labs(x="", y="", title="PACF For Avocado Average Price")
```

```{r}
# Forecast.
fit %>% forecast(h=27) %>% autoplot(level=NULL) + 
  theme_classic() + 
  autolayer(cali, average_price) + 
  labs(y="", title= "California's Forecasted Average Price Of Avocados",
       subtitle = "Jan 4, 2015 - Dec 2, 2018",
       x="")
```

## Ensembles

To create a combination of models we can use the fit object containing the mable. Specifically, any model in the fit object can be combined using a linear combination. The code below combines the AR(1) model and the LS model by finding a simple average between both models.

```{r}
fit %>%  mutate(Ensemble=0.5*LS+0.5*AR) -> fit2

#Now we can plot the forecasts for the ensemble, the AR1 and the LS model.

fit2 %>% select(AR,LS,Ensemble) %>%
  forecast(h=27) %>% autoplot(level=NULL) +
  autolayer(cali,average_price) + theme_classic() +
  labs(y="", title= "California's Forecasted Average Price Of Avocados",
       subtitle = "Jan 4, 2015 - Dec 2, 2018",
       x="")
```

```{r}
fit2 %>% forecast(h=27) %>% accuracy(cali)
```

## Exercises

### Question 1

Recall Question 2 in Module 6, where you are tasked to forecast the energy consumption of a new ice cream machine. Use all the data in the Electricity Bill file to estimate an AR1 model. Graph the ACF and PACF and confirm that an AR1 is appropriate. Check residuals from the model and confirm that they resemble white noise. Create a graph of the series and the forecast for four periods ahead.

```{r}
Elec<-read_csv("http://jagelves.github.io/Data/ElectricityBill.csv")

Elec %>% mutate(Date=yearmonth(paste(Year, Month, sep = "-"))) %>% 
  select(Date,`Bill Amount`) %>% 
  as_tsibble(index=Date)-> Elec_ts
```

```{r}
Elec_ts %>% ACF(`Bill Amount`,lag_max = 24) %>% autoplot()
Elec_ts %>% PACF(`Bill Amount`,lag_max = 24) %>% autoplot()
```

```{r}
Elec_ts %>% model(AR1=ARIMA(`Bill Amount`~pdq(1,0,0))) -> fit
```

```{r}
fit %>% augment() %>% ACF(.resid,lag_max = 24) %>% autoplot()
fit %>% augment() %>% PACF(.resid,lag_max = 24) %>% autoplot()
```

### Question 2

Recall Question 1 in Module 6, where you are tasked to forecast Tesla’s energy deployment. Use the entire data in the Electricity Bill Excel file to estimate an ensemble of the TSLM model with linear trend and seasonality, and a TSLM model with quadratic trend and seasonality. Let the weight of the linear model be 30% and that of the quadratic model 70%. Graph the forecast for the next four periods of the three models.

```{r}
library(ggthemes)
storage<-read_csv("http://jagelves.github.io/Data/teslaE.csv")

storage %>%
  mutate(Date=yearquarter(Date)) %>%
  as_tsibble(index=Date) -> storage_ts
```

```{r}
storage_ts %>%
  model(LM=TSLM(EnergyStorage~trend()+season()),
        LM2=TSLM(EnergyStorage~trend()+I(trend()^2)+season())) ->fit

fit %>% mutate(Ensemble=0.3*LM+0.7*LM2) -> fit2
```

```{r}
fit2 %>% forecast(h=4) %>% 
  autoplot(level=NULL) +
  autolayer(storage_ts,EnergyStorage) +
  theme_clean()
```

### Question 3

Using the monthly inflation data from the Inflation Excel file, forecast the inflation rate for the next four months. First, use the ARIMA() function to let the model automatically select the best-fitting ARIMA model for the inflation data while incorporating the interest rate into the model with the command ARIMA(Inflation~InterestRate). Develop two scenarios: the Federal Reserve holds the interest rate steady at 5.33% over the next four months, and the Fed follows a decreasing schedule of 5.25%, 5.00%, 4.75%, and 4.50%. Identify the month the Fed reaches its target of 2% inflation in each scenario and determine which schedule achieves this target faster. Create a graph that shows each scenario and the target inflation rate.

```{r}
library(ggthemes)
data<-read_csv("https://jagelves.github.io/Data/Inflation.csv")
```

```{r}
data %>% mutate(Date=yearmonth(mdy(Date))) %>% 
  as_tsibble(index=Date) -> data_ts
```

```{r}
data_ts %>% 
  model(ARIMA=ARIMA(Inflation~InterestRate)) -> fit2

Int <- scenarios(
  Keep = new_data(data_ts, 4) %>% 
    mutate(InterestRate=rep(5.33,4)),
  Lower = new_data(data_ts, 4) %>% 
    mutate(InterestRate=c(5.25,5.00,4.75,4.5)),
  names_to = "Scenario")
```

```{r}
forecast(fit2,new_data=Int)
```

```{r}
forecast(fit2,new_data=Int) %>% autoplot(level=95) + 
  autolayer(data_ts,Inflation) + 
  theme_clean() + 
  geom_hline(yintercept=2, lty=2) + 
  labs(title="US Monthly Inflation Rate",
       subtitle="Keep interest rate at 5.33% vs. 25 basis point reductions",
       x="",y="")
```

### Question 4

Rising U.S. household debt is concerning because it increases financial strain on families, reduces disposable income, and heightens vulnerability to interest rate hikes. It raises the risk of defaults and bankruptcies, potentially destabilizing financial institutions and the broader economy. Additionally, higher debt levels can slow consumer spending and economic growth, while also creating risks in the housing market, making the economy more fragile during downturns. Use the data in the U.S. Debt Excel file to forecast the US total debt of households using ARIMA. Start by graphing the ACF and PACF, then suggest two ARIMA models and let the algorithm choose the third. Calculate the BIC for the three models and select the model with the lowest BIC. Forecast the series into the future. By what date does the model predict the US Household debt will reach 20 trillion dollars?

```{r}
data<-read_csv("https://jagelves.github.io/Data/USDebt.csv")

data %>% mutate(Quarter=yearquarter(Quarter)) %>% 
  as_tsibble(index=Quarter) -> data_ts
```

```{r}
data_ts %>% ACF(diff(Total)) %>% autoplot()
data_ts %>% PACF(diff(Total)) %>% autoplot()
```

```{r}
data_ts %>% 
  model(AR3=ARIMA(Total~pdq(3,1,0)),
        AR1=ARIMA(Total~pdq(1,1,0)),
        ARIMA=ARIMA(Total)) -> fit

#Now we can observe the residuals:

fit %>% augment() %>% ACF(.resid) %>% autoplot()
```

```{r}
fit %>% glance() %>% select(.model,BIC)
```

```{r}
fit %>% forecast(h=20) %>% 
  as_tsibble() %>% filter(.model=="AR3") %>% 
  autoplot(.mean, col="blue", lty=2) +
  autolayer(data_ts,Total) + theme_clean() +
  labs(title="US Consumer Debt",
       subtitle="Trillion of Dollars",
       x="",y="") +
  geom_hline(yintercept=20,lty=2)
```

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
