---
title: "M4_Assignment"
author: "Danni Martin"
date: "11/18/2024"
format: pdf
editor: visual
---

## Q2 Home Prices

### Q2.1 Loading packages and data

Write code below that loads fpp3, tidyverse packages. Then load HomePrices.csv data set with the read_csv() function and save it in an object called Homes. Use the glimpse() function to preview the data. Include the code below:

```{r}
#| warning: false
library(tidyverse)
library(fpp3)

Homes <- read_csv("https://jagelves.github.io/Data/HomePrices.csv")

glimpse(Homes)
```

### Q2.3 Piping

Start with the Homes object. Use the pipe operator %\>% to chain the following commands.

-   filter the data so that only "two-bedroom" houses are included
-   coerce the saledate variable to a date using the dmy() function. format the saledate variable to quarters by using the yearquarter() function.
-   remove the bedrooms and type variables. You should include a pipe for each bullet point and save the result to a new object called h2.

```{r}
Homes %>%
  filter(bedrooms == 2, type == "house") %>%
  mutate(saledate = yearquarter(dmy(saledate))) %>%
  select(saledate, Price) -> h2

head(h2)
```

### Q2.4 tsibble

Coerce the h2 tibble to a tsibble using the as_tsibble() function. Make sure you specify the correct index. Save the new tsibble in an object called h2_ts. Include your code below:

```{r}
h2_ts <- as_tsibble(h2, index=saledate)
```

### Q2.5 Plot

Create a line plot of the Price variable using the ggplot() function. Add a title that reads "Quarterly Home Prices" and a subtitle that reads "2007 Q3 to 2019 Q3". Omit the x and y captions.

```{r}
ggplot(data=h2_ts) +
  geom_line(mapping = aes(x=saledate, y=Price), color='black') +
  theme_classic() +
  labs(x="",
       y="",
       title="Quarterly Home Prices",
       subtitle="2007 Q3 to 2019 Q3")
```

### Q2.6 STL

Perform an STL decomposition using the model() and STL() functions. Set the window argument to 11 for the trend(). Set 'periodic' for the window argument of the season() function.

```{r}
h2_ts %>%
  model(STL(Price~trend(window=11)+season(window='periodic'), robust=TRUE)) %>%
  components() %>% autoplot() + theme_classic()
```

### Q2.7 STL Components

```{r}
h2_ts %>%
  model(STL(Price~trend(window=11)+season(window='periodic'), robust=TRUE)) %>%
  components() -> h2_components
h2_components
```
