---
title: "M5_Assignment"
author: "Danni Martin"
date: "11/25/2024"
format: pdf
editor: visual
---

## Q2 Avocados

### Q2.1 Loading Packages and Data

Write code below that loads fpp3, tidyverse packages. Then load the avocado2020.csv data set using the read_csv() function and save it in an object called avocado.

```{r}
library(fpp3)
library(tidyverse)

avocado = read.csv("https://jagelves.github.io/Data/avocado2020.csv")
```

### Q2.2 Piping1

Use piping to accomplish the following tasks. Start with the avocado object. - remove the observations with a date of 1/1/18 using the filter() function. - use the mutate() and replace() functions to replace observations with a date of 1/7/19 to 1/6/19. - use the mutate() and replace() functions to replace observations with a date of 1/6/20 to 1/5/20. - filter the data so that the geography included is "Richmond/Norfolk" and the type of avocado is "organic". - select only the date and average_price variables Save the new data into an object called RVA. RVA should have 305 observations and two variables. Include your code below:

```{r}
avocado %>%
  filter(date!=("1/1/18")) %>%
  mutate(date = replace(date, date == ("1/7/19"), ("1/6/19"))) %>%
  mutate(date = replace(date, date == ("1/6/20"), ("1/5/20"))) %>%
  filter(geography == "Richmond/Norfolk", type == "organic") %>%
  select(date, average_price) -> RVA
```

### Q2.3 Piping2

Use piping to accomplish the following tasks. Start with the RVA object.

-   use the mutate() function to coerce the date variable to a date that has the year and week by nesting the mdy() function inside of the yearweek() function.
-   coerce the data to a tsibble using the as_tsibble() function.
-   filter the period "2019-01-07" to "2020-11-29" by using the filter_index() function. Save the result in an object called RVA_ts. RVA_ts should have 99 observations and two variables. Include your code below:

```{r}
RVA %>%
  mutate(date = yearweek(mdy(date))) %>%
  as_tsibble(index=date) %>%
  filter_index("2019-01-07"~"2020-11-29") -> RVA_ts
```

### Q2.4 Train and Test Sets

Create a test and training set and save them into objects called train_RVA and test_RVA\*, respectively. The training set should include the period 2019-01-06 to 2020-10-25, and the test set should include the period 2020-11-01 to 2020-11-29. Include the code below:

```{r}
RVA_ts %>%
  filter_index("2019-01-06"~"2020-10-25") -> train_RVA

RVA_ts %>%
  filter_index("2020-11-01"~"2020-11-29") -> test_RVA

```

### Q2.5 Autoplot

Create a plot of the average price of avocados (using RVA_ts) in Richmond/Norfolk using the autoplot() function. The title should read "Weekly Avocado Price in Richmond/Norfolk". The subtitle should read "2019 to 2020". Upload a screenshot of the plot below:

```{r}
RVA_ts %>%
  autoplot(average_price) + theme_classic() +
  labs(x = "",
       y= "",
       title = "Weekly Avocado Price in Richmond/Norfolk",
       subtitle = "2019 to 2020")
```

### Q2.6 Fit Models

Fit the following models using the model() and TSLM() functions. Use the training set.

LS: A simple linear model with the trend() parameter. LS3: A cubic model with the trend(), trend()\^2, and trend()\^3 parameters. LS6: A hexic model with the trend()...trend()\^6 parameters. Save the result in an object called fit. Include your code below:

```{r}
train_RVA %>%
  model(LS=TSLM(average_price~trend()),
        LS3=TSLM(average_price~trend()+I(trend()^2)+I(trend()^3)),
        LS6=TSLM(average_price~trend()+I(trend()^2)+I(trend()^3)
                 +I(trend()^4)+I(trend()^5)+I(trend()^6))) -> fit
```

### Q2.7 Plot1

Create a plot that shows the training set and the fitted values of all the models estimated (LS,LS3, and LS6). Use a red line for the LS model, an orange line for the LS3 model, and a blue line for the LS6 model.

```{r}
train_RVA %>% autoplot(average_price) + theme_classic() + 
  geom_line(aes(y = .fitted), col="red",
            data = augment(fit) %>% filter(`.model`=="LS")) +
geom_line(aes(y = .fitted), col="orange",
            data = augment(fit) %>% filter(`.model`=="LS3")) +
  geom_line(aes(y = .fitted), col="blue",
            data = augment(fit) %>% filter(`.model`=="LS6"))
```

### Q2.8 gt Package

Load the gt package. Use piping to accomplish the following tasks. Start with the fit object.

-   use the accuracy() function to find the model's fit for the training set.
-   select the variables .model and RMSE.
-   arrange the data using the arrange() function by the RMSE.
-   use the gt package to highlight the row (model) with the lowest RMSE.

```{r}
library(gt)

fit %>%
  accuracy() %>%
  select(.model, RMSE) %>%
  arrange(RMSE) %>%
  gt() %>%
  tab_style_body(
    style = cell_fill(color="lightgreen"),
    values = "LS6",
    targets ="row")
```

### Q2.9 Information Criterion

Use the fit object to retrieve the AIC, AICc, and BIC. According to the BIC, which model should be preferred? Why?

```{r}
# The LS6 Model as its value is lowest.
fit %>% glance() %>%
  select(".model", "AIC", "AICc", "BIC")
```

### Q2.10 CV

Use piping to perform cross-validation. Start with the train_RVA object.

-   use the stretch_tsibble() function with the argument .init set to 20 and the .step argument set to five.
-   estimate the LS, LS3 and LS6 models once again using the model() and TSLM() functions.
-   forecast five periods (h=5) using the forecast() function.
-   use the accuracy() function with the RVA_ts as only input.

```{r}
train_RVA %>%
  stretch_tsibble(.init = 20, .step = 5) %>%
  model(LS=TSLM(average_price~trend()),
        LS3=TSLM(average_price~trend()+I(trend()^2)+I(trend()^3)),
        LS6=TSLM(average_price~trend()+I(trend()^2)+I(trend()^3)
                 +I(trend()^4)+I(trend()^5)+I(trend()^6))) %>%
  forecast(h=5) %>%
  accuracy(RVA_ts)
```

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
