---
title: "M6_BUAD5122"
author: "Danni Martin"
date: "2/13/2024"
format: docx
editor: visual
---

## Problem 1

The problem is on support vector approach. The data set you will use to complete the problem is provided below.

In this problem, you will use support vector approaches in order to predict whether a given car gets high or low gas mileage based on the Auto data set.

```{r}
#Import libraries
library(ISLR)
library(e1071)
```

```{r}
#Load dataset
Auto <- ISLR::Auto
head(Auto)
```

### A

Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.

```{r}
#Create binary variable
Auto$mpg01 <- as.factor(as.numeric(Auto$mpg > median(Auto$mpg)))

#Verify
summary(Auto)
```

### B

Use the e1071 library in R. Use the SVM() function and kernel = "linear" and a cost argument to fit a support vector classifier to the data with three different values of cost at cost = .01, 0.1, 5 individually and respectively, in order to predict whether a car gets high or low gas mileage.  

```{r}
#Linear SVM with cost = .01
svm.01 <- svm(mpg01 ~., data = Auto, kernel = "linear", cost = .01)
print(svm.01)
```

```{r}
#Linear SVM with cost = .1
svm.1 <- svm(mpg01 ~., data = Auto, kernel = "linear", cost = .1)
print(svm.1)
```

```{r}
#Linear SVM with cost = 5
svm5 <- svm(mpg01 ~., data = Auto, kernel = "linear", cost = 5)
print(svm5)
```

### C

Use the tune() function and kernel = "linear" to report the cross-validation errors associated with different values of cost, use cost=c(0.01, 0.1, 5). Comment on your results.

```{r}
set.seed(1)
#Linear
tune_linear <- tune(svm, mpg01~., data = Auto, kernel = "linear", ranges = list(cost = c(.01, .1, 5)))
summary(tune_linear)
```

Cost = 5.0 results in the lowest cross-validation error rate (2.05%), followed by cost = .10 (4.6%), and lastly cost = .01 (7.65%). Fewer observations were misclassified as cost increases. We also observe fewer support vectors as cost increases due to smaller margins.

### D

Now repeat (c) with tune(). This time using SVMs with radial kernels and polynomial kernels. For radial kernels, use gamma=c(2, 5) and cost=c(0.01, 0.1, 5), For polynomial kernels, use degree=c(2, 3) and cost=c(0.01, 0.1, 5). Comment on your results.

```{r}
set.seed(1)
#Radial
tune_radial <- tune(svm, mpg01 ~., data=Auto, kernel = "radial", ranges = list(cost = c(.01, .1, 5), gamma = c(2, 5)))
summary(tune_radial)
```

Cost = 5 with gamma = 2 results in the lowest cross-validation error rate (13.5%), followed by cost = 5 with gamma = 5 (48.2%).
Once again higher cost results in fewer misclassified instances; however, using a radial kernel results in a worst fit when compared to the linear kernel.

```{r}
set.seed(1)
#Polynomial
tune_poly <- tune(svm, mpg01 ~., data=Auto, kernel = "polynomial", ranges = list(cost = c(.01, .1, 5), degree = c(2, 3)))
summary(tune_poly)

```
Using a polynomial kernel results in a cross-validation error rate of 55.1%, regardless of degree or cost. This suggests a very poor fit.

The linear kernel performed better than the radial and polynomial kernels. This suggests that the data is linearly separable. The linear kernel model with cost = 5 performed the best overall. 


```{r}

```


